---
title: "Final Project"
author: "Zane Brotherton, Nate Adam, Feolu Kolawole"
date: "2023-08-14"
output: pdf_document
---

# Project Description

```{r}
set.seed(123)
House_Data = read.table(url('https://raw.githubusercontent.com/zbrotherton2158/okadaStats/master/ames2000_NAfix.csv'), sep=',', header=T)
```

## The study
questions to be answered are the following:
• What field does the data come from?
  This data comes from the assessors office of Ames, Iowa
• What are the goals of the study? Are there any effects of particular interest?
  The purpose of this data is to be able to predict the value of a house in Ames Iowa
  Our areas of interest are the effect of the neighborhoods, lot area, total rooms above grade, year built, and total square feet on the price of the house. 
• How might these goals be answered, i.e. tests / confidence intervals?
  This goal could be answered by a prediction model that, given information about a house, will output a 95% confidence interval predicting the price of the house.

## The data
### Predictor Variable Spread
```{r}
require(tidyverse)
Numerical_Count = length(select_if(House_Data, is.numeric))
Categorical_Count = length(select_if(House_Data, negate(is.numeric)))

```

*Clean "none" values*
```{r}

# NOTE: Might have to revamp
# Turned None's into 0's for respective rows

#for(colName in colnames(House_Data)){
#  if(colName != "Garage.Yr.Blt"){
#    print(colName)
#    colName = replace(colName, colName == ("None"), 0)
#  }
#}

# j is the column of each selected box
for (j in 1:80){ 
  # clear the haveNumbers in case it has values
  haveNumbers = c()
  # i is the row of each selected box
  for (i in 1:2000) {
    # if the selected box is empty, fill it with "None"
    if (House_Data[i,j] == "") {
      House_Data[i,j] = "None"
    }
    # record whether the selected box contained a number in a list called haveNumbers
    haveNumbers = c(haveNumbers, !grepl("\\D", House_Data[i,j]))
  }
  # if any of the selected boxes in the column contain a number,
  # replace all "None"s in the column with 0s
  if (any(haveNumbers) && (j != 59)) { 
    House_Data[j] = replace(House_Data[j], House_Data[j] == "None", 0)
  }
}
# print the data set to see if it worked
print(House_Data)

#House_Data = replace(House_Data, House_Data == 'None', 0)
#print(as.numeric(House_Data$Street))
#print(data.frame(select_if(House_Data, negate(is.numeric))))
```

*Partition the data*
```{r}
N = nrow(House_Data)
num_train = as.integer(0.7 * N)
train_idx = sample(1:N, size=num_train, replace=F)
test_idx = setdiff(1:N, train_idx)
House_Data_Train = House_Data[train_idx,]
House_Data_Test = House_Data[test_idx,]
```

### Correlation Info
*Numerical Correlation*
```{r}
require(corrplot)
require(dplyr)
```
Numerical_Correlations = cor(House_Data_Numerical, House_Data_Numerical$SalePrice)

### Analysis
• How are the predictor variables spread out? Are there any noteworthy features to their spread
that could be highly influential observations?

The variables are seemingly random between discrete, ordinal, nominal, and continuous types of variables. Of these, we found 26 to be numerical and 54 categorical. Within the discrete variables, variables with a spread like fireplaces where despite being numerical only have values of 1 - arbitrarily small number may have a higher effect on the price per unit increase. 

• Are any of the predictor variables highly correlated?

  The highest correlation numerical variables are:
  
  • Overall Quality (~0.805)
  
  • Gr. Living Area (~0.720)
  
  • Year Built (~0.57)
  
  • Full Bath (~0.561)
  
  • Year Remodeled (~0.535)
  
  • Rooms Above Ground (~0.504)
  
  • Fireplace (~0.48)
  
  • 1st Floor Square Ft. (~0.619)
  
  • 2nd Floor Square Ft. (~0.295)*
  

## The models 
• Which predictor variables, if any, should be included in the model a priori?
  neighborhoods, lot area, total rooms above grade, year built, and total square feet
• Are there any interactions that should be considered for inclusion in the model?
  
• Are there any three way interactions that should be considered?
• Are there any interactions that should NOT be considered?

Model Draft
```{r}
#factorize these?
model = lm(sqrt(SalePrice) ~ 
             #Numerical
             Overall.Qual +
             Gr.Liv.Area +
             Year.Built +
             Year.Remod.Add +
             Full.Bath +
             #Categorical
             Neighborhood +
             Condition.1 +
          #   Condition.2 +
             Bldg.Type +
             House.Style +
             Roof.Matl +
             Exter.Qual +
             BsmtFin.Type.1 +
             Kitchen.Qual +
             Fireplace.Qu +
             Garage.Type +
             Pool.QC
           , House_Data_Train)
summary(model)
plot(model)

```
```{r}
#to investigate numerical variables
House_Data_Numerical = data.frame(select_if(House_Data, is.numeric))

print(cor(House_Data_Numerical, House_Data_Numerical$SalePrice))
```

```{r}
#to investigate categories
library(tidyverse)
House_Data_Categorical = data.frame(select_if(House_Data, negate(is.numeric)))
House_Data_Categorical$SalePrice = House_Data$SalePrice

boxplot(House_Data$SalePrice ~ House_Data$Full.Bath)

sapply(lapply(House_Data_Categorical, unique), length)

investigative_lm = lm(SalePrice ~ ., data = subset(House_Data_Categorical, 
select = -c(Utilities,Lot.Frontage,Mas.Vnr.Area,BsmtFin.SF.1,BsmtFin.SF.2,Bsmt.Unf.SF,Total.Bsmt.SF,Garage.Yr.Blt,Garage.Area)))
summary(investigative_lm)
```

R^2 of testing set
```{r}
plot(House_Data_Test$SalePrice, predict(model, newdata = House_Data_Test, type = "response"))


r2 = 1 - (sum((House_Data_Test$SalePrice - predict(model, newdata = House_Data_Test, type = "response")) ^ 2)
          /sum((House_Data_Test$SalePrice - mean(House_Data_Test$SalePrice)) ^ 2))
print(r2)
```

*Results*: In this section, you should report the results obtained by fitting the proposed models in the
previous section. Emphasis should be placed on clarity, as if the report were a statistical consultant’s
report for a nonstatistician. For instance, loads of R output would, in general, not be acceptable. Plots
and well-organized tables are good things to have in this section. Possible questions to be addressed
here are the following:
• What is the final regression model for the data?
• Using the standard diagnostic tests, does the model appear to fit the data well?
• What are the final confidence intervals for the effects of interest mentioned in the study section?
Use your validation data to construct these intervals. Do these intervals seem very sensitive to the
choice of model (i.e. do they vary widely for different choices of variables in the model)?
• What is your estimated prediction accuracy for your model? (Evaluated on the validation set).
• Compare the intervals constructed using your final selected model fit to the validation set to the
same intervals constructed on the training set. Are they very different? Which do you believe
more?

## Appendix 
Ideally, there will be comments in the file, i.e. lines beginning with “#” to clarify what each part of the code is doing.

## Acknowledgements : If you consult outside sources that refer to this data set, you should cite these
as references, and describe what you used from each source. Sources include material found on the
internet, journal articles and books.
