---
title: "Final Project"
author: "Zane Brotherton, Nate Adam, Feolu Kolawole"
date: "2023-08-14"
output: pdf_document
---

# Project Description

```{r}
set.seed(123)
House_Data = read.table(url('https://raw.githubusercontent.com/zbrotherton2158/okadaStats/master/ames2000_NAfix.csv'), sep=',', header=T)
```

## The study
questions to be answered are the following:
• What field does the data come from?
  This data comes from the assessors office of Ames, Iowa
• What are the goals of the study? Are there any effects of particular interest?
  The purpose of this data is to be able to predict the value of a house in Ames Iowa
  Our areas of interest are the effect of the neighborhoods, lot area, total rooms above grade, year built, and total square feet on the price of the house. 
• How might these goals be answered, i.e. tests / confidence intervals?
  This goal could be answered by a prediction model that, given information about a house, will output a 95% confidence interval predicting the price of the house.

## The data
### Predictor Variable Spread
```{r}
require(tidyverse)
Numerical_Count = length(select_if(House_Data, is.numeric))
Categorical_Count = length(select_if(House_Data, negate(is.numeric)))

```

*Clean "none" values*
```{r}

# NOTE: Might have to revamp
# Turned Nones into 0's for respective rows
# Filled Empty boxes with "None" for categorical variables. 

for (j in 1:80){ 
  haveNumbers = c()
  for (i in 1:2000) {
    if (House_Data[i,j] == "") {
      House_Data[i,j] = "None"
    }
    haveNumbers = c(haveNumbers, !grepl("\\D", House_Data[i,j]))
  }
  if (any(haveNumbers) && (j != 59)) { 
    House_Data[j] = replace(House_Data[j], House_Data[j] == "None", 0)
  }
}

```

*Partition the data*
```{r}
N = nrow(House_Data)
num_train = as.integer(0.7 * N)
train_idx = sample(1:N, size=num_train, replace=F)
test_idx = setdiff(1:N, train_idx)
House_Data_Train = House_Data[train_idx,]
House_Data_Test = House_Data[test_idx,]
```

### Correlation Info
*Numerical Correlation*
```{r}
require(corrplot)
require(dplyr)
```
Numerical_Correlations = cor(House_Data_Numerical, House_Data_Numerical$SalePrice)

### Analysis
• How are the predictor variables spread out? Are there any noteworthy features to their spread
that could be highly influential observations?

The variables are seemingly random between discrete, ordinal, nominal, and continuous types of variables. Of these, we found 26 to be numerical and 54 categorical. Within the discrete variables, variables with a spread like fireplaces where despite being numerical only have values of 1 - arbitrarily small number may have a higher effect on the price per unit increase. 

• Are any of the predictor variables highly correlated?

  The highest correlation numerical variables are:
  
  • Overall Quality (~0.805)
  
  • Gr. Living Area (~0.720)
  
  • Year Built (~0.57)
  
  • Full Bath (~0.561)
  
  • Year Remodeled (~0.535)
  
  • Rooms Above Ground (~0.504)
  
  • Fireplace (~0.48)
  
  • 1st Floor Square Ft. (~0.619)
  
  • 2nd Floor Square Ft. (~0.295)*
  
  The most statistically significant categorical variables are:
  
  

## The models 

Model Draft
```{r}
#factorize these?
model = lm(SalePrice ~ 
             #Numerical
             Overall.Qual +
             Gr.Liv.Area +
             Year.Built +
             Year.Remod.Add +
             Full.Bath +
             #Categorical
             Neighborhood +
             Condition.1 +
             Condition.2 +
             Bldg.Type +
             House.Style +
             Roof.Matl +
             Exter.Qual +
             BsmtFin.Type.1 +
             Kitchen.Qual +
             Fireplace.Qu +
             Garage.Type +
             Pool.QC
           , House_Data)
summary(model)
plot(model)

```
```{r}
#to investigate numerical variables
House_Data_Numerical = data.frame(select_if(House_Data, is.numeric))

print(cor(House_Data_Numerical, House_Data_Numerical$SalePrice))
```

```{r}
#to investigate categories
library(tidyverse)
House_Data_Categorical = data.frame(select_if(House_Data, negate(is.numeric)))
House_Data_Categorical$SalePrice = House_Data$SalePrice

boxplot(House_Data$SalePrice ~ House_Data$Full.Bath)

sapply(lapply(House_Data_Categorical, unique), length)

investigative_lm = lm(SalePrice ~ ., data = subset(House_Data_Categorical, 
select = -c(Utilities,Lot.Frontage,Mas.Vnr.Area,BsmtFin.SF.1,BsmtFin.SF.2,Bsmt.Unf.SF,Total.Bsmt.SF,Garage.Yr.Blt,Garage.Area)))
summary(investigative_lm)
```

### Analysis

• Which predictor variables, if any, should be included in the model a priority?

The highest correlation numerical variables and most statistically significant categorical listed under "The Data" are of the most interest and are the most prioritized in the model. (Neighborhoods, lot area, total rooms above grade, year built, and total square feet etc.)

• Are there any interactions that should be considered for inclusion in the model?

• Are there any three way interactions that should be considered?

• Are there any interactions that should NOT be considered?


## Results

• What is the final regression model for the data?

• Using the standard diagnostic tests, does the model appear to fit the data well?

• What are the final confidence intervals for the effects of interest mentioned in the study section?

Use your validation data to construct these intervals. Do these intervals seem very sensitive to the
choice of model (i.e. do they vary widely for different choices of variables in the model)?
• What is your estimated prediction accuracy for your model? (Evaluated on the validation set).
• Compare the intervals constructed using your final selected model fit to the validation set to the
same intervals constructed on the training set. Are they very different? Which do you believe
more?

## Appendix 
Ideally, there will be comments in the file, i.e. lines beginning with “#” to clarify what each part of the code is doing.

## Acknowledgements : If you consult outside sources that refer to this data set, you should cite these
as references, and describe what you used from each source. Sources include material found on the
internet, journal articles and books.
