---
title: "Final Project"
author: "Zane Brotherton, Nate Adam, Feolu Kolawole"
date: "2023-08-14"
output: pdf_document
---

**Project Description**:

```{r}
set.seed(123)
House_Data = read.table(url('https://raw.githubusercontent.com/zbrotherton2158/okadaStats/master/ames2000_NAfix.csv'), sep=',', header=T)
```

*The study*: In this section, you should give a description of the study underlying the dataset. Possible
questions to be answered are the following:
• What field does the data come from?
  This data comes from the assessors office of Iowa
• What are the goals of the study? Are there any effects of particular interest?
  The purpose of this data is to be able to predict the value of a house in Ames Iowa
  Our areas of interest are the effect of the neighborhoods, lot area, total rooms above grade, year built, and total
  square feet on the price of the house. 
• How might these goals be answered, i.e. tests / confidence intervals?
  

*The data*: In this section, you should describe the data set and possibly do some exploratory data
analysis. For instance:
• How are the predictor variables spread out? Are there any noteworthy features to their spread
that could be highly influential observations?
• Are any of the predictor variables highly correlated?

```{r}
require(corrplot)
require(dplyr)

```

*The models*: In this section, you should develop a model for the data that will allow them to answer
some of the specific goals of the study. Possible questions to be addressed here are the following:
• Which predictor variables, if any, should be included in the model a priori?
• Are there any interactions that should be considered for inclusion in the model?
• Are there any three way interactions that should be considered?
• Are there any interactions that should NOT be considered?

Model Draft
```{r}
#factorize these?
model = lm(SalePrice ~ 
             #Numerical
             Overall.Qual +
             Gr.Liv.Area +
             Year.Built +
             Year.Remod.Add +
             Full.Bath +
             #Categorical
             Neighborhood +
             Condition.1 +
             Condition.2 +
             Bldg.Type +
             House.Style +
             Roof.Matl +
             Exter.Qual +
             BsmtFin.Type.1 +
             Kitchen.Qual +
             Fireplace.Qu +
             Garage.Type +
             Pool.QC
           , House_Data)
summary(model)
plot(model)

```
```{r}
#to investigate numerical variables
House_Data_Numerical = data.frame(select_if(House_Data, is.numeric))

print(cor(House_Data_Numerical, House_Data_Numerical$SalePrice))
```

```{r}
#to investigate categories
library(tidyverse)
House_Data_Categorical = data.frame(select_if(House_Data, negate(is.numeric)))
House_Data_Categorical$SalePrice = House_Data$SalePrice

boxplot(House_Data$SalePrice ~ House_Data$Full.Bath)

sapply(lapply(House_Data_Categorical, unique), length)

investigative_lm = lm(SalePrice ~ ., data = subset(House_Data_Categorical, 
select = -c(Utilities,Lot.Frontage,Mas.Vnr.Area,BsmtFin.SF.1,BsmtFin.SF.2,Bsmt.Unf.SF,Total.Bsmt.SF,Garage.Yr.Blt,Garage.Area)))
summary(investigative_lm)
```


*Results*: In this section, you should report the results obtained by fitting the proposed models in the
previous section. Emphasis should be placed on clarity, as if the report were a statistical consultant’s
report for a nonstatistician. For instance, loads of R output would, in general, not be acceptable. Plots
and well-organized tables are good things to have in this section. Possible questions to be addressed
here are the following:
• What is the final regression model for the data?
• Using the standard diagnostic tests, does the model appear to fit the data well?
• What are the final confidence intervals for the effects of interest mentioned in the study section?
Use your validation data to construct these intervals. Do these intervals seem very sensitive to the
choice of model (i.e. do they vary widely for different choices of variables in the model)?
• What is your estimated prediction accuracy for your model? (Evaluated on the validation set).
• Compare the intervals constructed using your final selected model fit to the validation set to the
same intervals constructed on the training set. Are they very different? Which do you believe
more?

*Appendix*: In this section, you should attach a final, editied, copy of the R code used in the analysis.
Ideally, there will be comments in the file, i.e. lines beginning with “#” to clarify what each part of the code is doing.

*Acknowledgements*: If you consult outside sources that refer to this data set, you should cite these
as references, and describe what you used from each source. Sources include material found on the
internet, journal articles and books.
